---
title: "Reverse Engineering Project"
author: "James Hartner, Daniel Lawall, Daniel Philipose "
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

Our data analysis project reverse engineers five major data findings from a Washington Post article titled “Nearly 250 women have been fatally shot by police since 2015,” which was published a few months after Breonna Taylor was shot and killed by police. The story is part of a bigger, ongoing project of the Washington Post; they have been collecting data on every deadly police shooting in the United States since January 2015. This database is still updated regularly, which explains why the findings of each of our “reverse engineering tasks” yielded results that don’t exactly equal the findings in the article (the article was published in September 2020, while the data we used is from the database last updated on September 28, 2021). 

According to the Post, which publishes some of the broader findings from their database on a page called “Fatal Force,” their data collection process “relies primarily on news accounts, social media postings and police reports.”

The main finding of the article was that Black women are killed by police at a disproportionate rate; of the 247 women killed in a police shooting since 2015 (and until the time the article was published), 48 were black. So, roughly 20% of the victims of all female, fatal police shootings were Black, while Black women only make up 13% of the U.S. female population. This particular data finding is reverse engineered below. 

Overall, the reporter(s) who worked on this story did a good job of using the data available to produce a story with depth. They also did a good job of “bringing the data to life”; the reporter elaborated on a few specific instances of fatal police shootings, providing more details and getting quotes from loved ones of the victims. Doing so helps to humanize the story. This story effectively connected Breonna Taylor’s death to a larger web of Black women who were killed by police. 

The reporters used three graphs/data visualizations total in the story, and each one was visually appealing and communicated important data findings. However, they could have used some more data visualizations, as the other major data findings were simply posted in-text. Additionally, this article doesn’t show how they came up with their findings, although they do include a link to the “Fatal Force” page, which does offer some more insight into the article’s findings (it has a “methodology” section that delves deeper into the Post’s process of collecting the data and putting the database together). 

The linked “Fatal Force” page also includes a link to the Washington Post GitHub, where the database is available for download. The database is well-organized and includes important characteristics of each case of a fatal police shooting (name of victim, age, race, gender, etc.). The reporter also does not communicate how the data may be flawed; for example, upon parsing through the data ourselves, there are cases of “N/A” being entered in certain categories for some data entries (examples when race, gender, etc. could not be confirmed). The Post created this database themselves, and it likely required lots of data cleaning and verification before the data was ready to be published. 


## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
Washington_Post_data <-read_csv("data/fatal-police-shootings-data.csv")

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: ["About 31 percent, or 77, of the 247 women fatally shot by police since 2015 had mental health issues, compared to 22 percent of the 5,362 men killed."]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?
Our findings were a little different because the story was posted over a year ago (Sept. 4, 2020) and the Washington Post have keep updating the data. In the data, which was taken Sept. 28, 2020, we found that since 2015, the percentage of women fatally shot by police who had mental health issues was higher at roughly 32 percent (95 of the 293 women). Men who had mental health issues was around 22 percent still ( 1398 of the 6303 men killed).]

```{r}

# Put code to reverse engineer sentence here

  
mental_illness <- Washington_Post_data %>%
  #filter(gender=="F") %>%
  group_by(gender,signs_of_mental_illness) %>%
  summarise(
    total=n()
  ) %>%
  mutate(percentage=total/sum(total)) %>%
  arrange(desc(total))


# Put code to reverse engineer sentence here
women_fatally_shot <- Washington_Post_data %>%
  filter(gender=="F") %>%
  group_by(signs_of_mental_illness) %>%
  summarise(
    total=n()
  ) %>%
  mutate(percentage=total/sum(total)) %>%
  arrange(desc(total))

#Men percentage
men_fatally_shot <- Washington_Post_data %>%
  filter(gender =="M") %>%
  group_by(signs_of_mental_illness) %>%
  summarize(
    total=n(),
    ) %>%
    mutate(percentage=total/sum(total)) %>%
  arrange(desc(total))
# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known. But within this small subset, Black women, who are 13 percent of the female population, account for 20 percent of the women shot and killed and 28 percent of the unarmed deaths]
* **Analysis summary**: [We broke this statement into three parts in order to determine each of the listed  facts. The process was similar across all three parts, in which we either filtered or grouped by race and gender. In each part, we also used "mutate" to create a new column to determine the percentages. We were able to confirm all three of the facts in this statement: Black women account for less than 1% of fatal shootings where race is know, 20% of the women shot and 28% of the total number of unarmed deaths.]

```{r}
# Put code to reverse engineer sentence here

overall_black_women <- Washington_Post_data %>%
  group_by(race, gender) %>%
   summarise(
    count=n()
  ) %>%
  mutate(pct=(count/5841)*100)
  

Black_women <- Washington_Post_data %>%
  filter(gender == "F") %>%
  group_by(race) %>%
  summarise(
    count=n()
  ) %>%
  mutate(pct_by_race=count/sum(count))

Unarmed_black_women <- Washington_Post_data %>%
  filter(gender == "F") %>%
  filter(armed == "unarmed") %>%
  group_by(race) %>%
  summarise(
    count=n()
  ) %>%
  mutate(pct_by_race=count/28)
  
# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [The starkest difference between women and men is the rate: Women account for about half the population, but 4 percent of the killings. Of those fatally shot every year, about 44 have been women.]
* **Analysis summary**: []

```{r}
# Put code to reverse engineer sentence here
Washington_Post_data %>%
filter(gender=="F")

women_fatally_shot <- Washington_Post_data %>%
  filter(gender=="F") %>%
  summarise(
   count = n()
  ) %>%
mutate(pct=(count/5))

women_fatally_shot %>%
  summarise(
   count = n()
  ) %>%
  mutate(pct=(count/6594))



# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: ["The starkest difference between women and men is the rate: Women account for about half the population, but 4 percent of the killings. Of those fatally shot every year, about 44 have been women."]
* **Analysis summary**: [This data was able to reasonably confirm the statistics used in the story, with slight changes due to unavailable variable values.]

```{r}
# Put code to reverse engineer sentence here
Washington_Post_data %>%
filter(gender=="F")

women_fatally_shot <- Washington_Post_data %>%
  filter(gender=="F") %>%
  summarise(
   count = n()
  ) %>%
mutate(pct=(count/5))

women_fatally_shot %>%
  summarise(
   count = n()
  ) %>%
  mutate(pct=(count/6594))

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [“The average age of women and men killed by police was 37, and about one-third were 25 to 34.”]
* **Analysis summary**: [The numbers found were able to reasonably confirm the data used in the report. The slight margin for error came within age values that had either been redacted or removed, which changed the final results down to decimal places but no further.]

```{r}
# Put code to reverse engineer sentence here

#Average for women
women_fatally_shot <- Washington_Post_data %>%
  filter(gender=="F")

age_women_fatally_shot <- women_fatally_shot %>%
  drop_na() %>%
  summarise(
    mean_age = mean(age)
  )

# Average for men
men_fatally_shot <- Washington_Post_data %>%
  filter(gender=="M")

age_men_fatally_shot <- men_fatally_shot %>%
   drop_na() %>%
  summarise(
    mean_age = mean(age)
  )


#one-third 25 to 34
Washington_Post_data %>%
  filter(age >= "25" & age <= "34") %>%
  summarise(
    spec_count = n()
  ) %>%
mutate(pct=(spec_count/6594)*100)

Washington_Post_data %>%
  summarise(
    total_count = n()
  ) 

  
# Display results of code below this codeblock
glimpse(age_men_fatally_shot)
glimpse(age_women_fatally_shot)

# Display results of code below this codeblock

```

-30-